# Topic 7 Exercises: Nonlinear regression

# AJ Imholte

#Theory Questions

7.9.3 -
```{r}
x = -2:2
y = 1 + x + -2 * (x-1)^2 * I(x>1)
plot(x,y)
```
The curve is linear between -2 and 1 (y = 1 + x) an quadratic between 1 and 2 (y = 1 + x - 2(x-1)^2).

7.9.4-
```{r}
x = -2:2
y = c(1 + 0 + 0, #x=2
      1 + 0 + 0, #x=-1
      1 + 1 + 0, #x=0
      1 + (1-0) + 0, #x=1
      1 + (1-1) + 0 #x=2)
)
plot(x,y)
```
The curve is constance between -2 and 0 (y = 1) and constant between 0 and 1 (y = 2) and linear between 1 and 2 (y = 3 - x).

7.9.5-

a. As lambda approaches infinity, the smoothing spline \hat({g_2}) will mostlikely have the smaller training RSS because it will be a higher-order polynomial due to the power of the penalty term. This will result in the model being more flexible.

b. As lambda approaches infinity, we would expect that \hat({g_1}) would have the smaller test RSS. This is due to \hat({g_2})'s degree of flexibility, which could possibly lead to overfitting.

c. When lambda is equal to zero, \hat({g_1}) and \hat({g_2}) are equivalent, so they will both have the same training and test RSS.

#Applied

7.9.11

a.
```{r}
set.seed(1)
x1 = rnorm(100)
x2 = rnorm(100)
e = rnorm(100, sd = 0.1)
Y = -5 + 1.5 * x1 + -2 * x2 + e
```

b.
```{r}
beta_0 = rep(NA, 1000)
beta_1 = rep(NA, 1000)
beta_2 = rep(NA, 1000)
beta_1[1] = 5
```

c.
```{r}
for (i in 1:1000) {
  a = Y - beta_1[i] * x1
  beta_2[i] = lm(a ~ x2)$coef[2]
  a = Y - beta_2[i] * x2
  lm_fit = lm(a ~ x1)
  if (i < 1000) {
    beta_1[i + 1] = lm_fit$coef[2]
  }
  beta_0[i] = lm_fit$coef[1]
}
plot(1:1000, beta_0, type = "l", xlab = "Iteration", ylab = "betas", ylim = c(-5, 6), col = "blue")
lines(1:1000, beta_1, col = "red")
lines(1:1000, beta_2, col = "purple")
legend("center", c("beta_0", "beta_1", "beta_2"), lty = 1, col = c("blue", "red", "purple"))
```
Coefficients quickly attain their least square values as the number of interations increase

f.
```{r}
lm_fit = lm(Y ~ x1 + x2)
plot(1:1000, beta_0, type = "l", xlab = "Iteration", ylab = "Betas", ylim = c(-6, 6), col = "blue")
lines(1:1000, beta_1, col = "red")
lines(1:1000, beta_2, col = "purple")
abline(h = lm_fit$coef[1], lty = "dashed", lwd = 3, col = rgb(0,0,0, alpha = 0.4))
abline(h = lm_fit$coef[2], lty = "dashed", lwd = 3, col = rgb(0,0,0, alpha = 0.4))
abline(h = lm_fit$coef[3], lty = "dashed", lwd = 3, col = rgb(0,0,0, alpha = 0.4))
legend("center", c("beta_0", "beta_1", "beta_2", "multiple regression"), lty = c(1, 1, 1, 2), col = c("blue", "red", "purple"))
```
The dotted lines on the graph above demonstrate that the multiple regression coefficients match exactly with the coefficients obtained using backfitting.

g. Based on the graphs above, when the relationship between Y and X's is linear, one iteration is enough to attain an accurate approximation of true regression coefficients.

